Estimate of the Neural Network Dimension Using Algebraic Topology and Lie Theory\thanks{The code can be found at: \href{https://github.com/karhunenloeve/neural-topology-lie}{https://github.com/karhunenloeve/neural-topology-lie}.}

### 2020-07-20 llncs2e/paper12.tex
typo notation
papers

A metric describes the geometric properties of a manifold. The change of coordinate systems and the learned metric tensor of a smooth manifold during backpropagation was formalized \cite{HauserR17}, which shows, that we operate often on a model of a Lie Group. Further, it has been shown that dense neural networks cannot approximate all functions arbitrarily precisely \cite{Johnson19}. Thereupon, neural networks with residual connections were investigated. Residual networks add the output of one layer to a deeper one, bypassing some of the layers in between. Indeed, they are universal function approximators \cite{LinJ18}. Inspired by finite difference methods, such a residual layer was defined as a forward or backward difference operator on a partition of layers. The state space dimension with residual connections is homeomorphic to $\mathbb{R}^{k \cdot n}$, where $n = \text{dim} \; M$ and $k$ is the number of times the difference operator was used \cite{HauserGJR19}. Such a layer is able to embed in $k\cdot n$ dimensional Euclidean space. The smooth manifold perspective inspired us to investigate its invariants leading into the theory discussed in this paper.

@@

We count the representatives of homology classes in the persistence landscapes to derive the dimension of the sought manifold. We get for $\mathcal{S}^1$:
\begin{align}
    H_0(\mathcal{S}^1) &= H_1(\mathcal{S}^1) = \mathbb{Z},\\
    H_i(\mathcal{S}^1) &= 0, \; \text{for all } i \geq 2.
\end{align}
Note, in Eq. \ref{homologykunneth} terms remain only for indices $i_{j} \in \{0,1\}$. Thus, we get
\begin{align}
    H_0(\mathbb{R}^p) &\simeq\mathbb{Z}^{p},\\
    \label{determineq}
    H_k(\mathbb{T}^{q}) &\simeq H_k(\mathcal{S}^{1}_{1} \times \cdots \times \mathcal{S}^{1}_{q}) \simeq \mathbb{Z}^{{q\choose{k}}},
\end{align}
where $p$ indicates the number of connected components. The number of local maxima for a homology group in the persistence landscape is the cardinality of the set of points $(\epsilon_i,\epsilon_{j+1})$ with derivatives $d\lambda^{\epsilon_i}_{\epsilon_{j+1}}(x) = 0 \; \text{and} \; d^2 \lambda^{\epsilon_i}_{\epsilon_{j+1}}(x) < 0$, for an example of these maximum values within a persistence landscape see Fig. \ref{figure2} b) c). Also, it is the solution for the binomial coefficient:
\begin{align}
     {q \choose k} = \frac{q}{1} \cdot \frac{q-1}{2} \cdots \frac{q-(k-1)}{k} = \prod_{i=1}^{k} \frac{q+1-i}{i}.
\end{align} 
We count in $PL(X)$ the elements of the $0$th homology group. The amount of elements from higher homology groups correspond to $q$ (Eq. \ref{determineq}) and are computed using Monte-Carlo methods for an (approximate) integer solution, see Tab. \ref{figure3}.

>>>>

A metric describes the geometric properties of a manifold. The change of coordinate systems and the learned metric tensor of a smooth manifold during backpropagation was formalized \cite{HauserR17}, which shows, that we operate often on a model of a Lie group. Further, it has been shown that dense neural networks cannot approximate all functions arbitrarily precisely \cite{Johnson19}. Thereupon, neural networks with residual connections were investigated. Residual networks add the output of one layer to a deeper one, bypassing some of the layers in between. Indeed, they are universal function approximators \cite{LinJ18}. Inspired by finite difference methods, such a residual layer was defined as a forward or backward difference operator on a partition of layers. The state space dimension with residual connections is homeomorphic to $\mathbb{R}^{k \cdot n}$, where $n = \text{dim} \; M$ and $k$ is the number of times the difference operator was used \cite{HauserGJR19}. Such a layer is able to embed in $k\cdot n$ dimensional Euclidean space. The smooth manifold perspective inspired us to investigate its invariants leading into the theory discussed in this paper.

@@

We count the representatives of homology classes in the persistence landscapes to derive the dimension of the sought manifold. We get for $\mathcal{S}^1$:
\begin{align}
    H_0(\mathcal{S}^1) &= H_1(\mathcal{S}^1) = \mathbb{Z},\\
    H_i(\mathcal{S}^1) &= 0, \; \text{for all } i \geq 2.
\end{align}
Note, in Eq. \ref{homologykunneth} terms remain only for indices $i_{j} \in \{0,1\}$. Thus, we get
\begin{align}
    H_0(\mathbb{R}^p) &\simeq\mathbb{Z}^{p},\\
    \label{determineq}
    H_k(\mathbb{T}^{q}) &\simeq H_k(\mathcal{S}^{1}_{1} \times \cdots \times \mathcal{S}^{1}_{q}) \simeq \mathbb{Z}^{{q\choose{k}}},
\end{align}
where $p$ indicates the number of connected components. The number of local maxima for a homology group in the persistence landscape is the cardinality of the set of points $(\epsilon_i,\epsilon_{j+1})$ with derivatives $d\lambda^{\epsilon_i}_{\epsilon_{j+1}}(x) = 0 \; \text{and} \; d^2 \lambda^{\epsilon_i}_{\epsilon_{j+1}}(x) < 0$, for an example of these maximum values within a persistence landscape see Fig. \ref{figure2} b) c). Also, it is the solution for the binomial coefficient:
\begin{align}
     {q \choose k} = \frac{q}{1} \cdot \frac{q-1}{2} \cdots \frac{q-(k-1)}{k} = \prod_{i=1}^{k} \frac{q+1-i}{i}.
\end{align} 
We count in $\text{Pl}(X)$ the elements of the $0$th homology group. The amount of elements from higher homology groups correspond to $q$ (Eq. \ref{determineq}) and are computed using Monte-Carlo methods for an (approximate) integer solution, see Tab. \ref{figure3}.
